Poker Chips, Earthquakes, and Computation

### 2014-10-11 finaldraft.tex
finished section on definitions on notation. still need to add figures and figure references though

We can coordinatize the grid graph by labeling it like the standard Cartesian grid, where we fix one node to be the origin $(0,0)$, and we index all other nodes by integer coordinates $(i,j)$. We will define the board as a function from the Cartesian grid to the nonnegative integers.

@@

This function maps node labels given as integer coordinates to chip stack sizes. For example, if a board $b$ has 15 chips at the node with coordinate $(5,4)$, then $b(5,4) = 15$.

@@

Recall from the introduction that an earthquake (abbreviated as $E$) is an operation that hits one node with more than $4$ chips and redistributes 1 chip to each neighbor. We define this operation in terms of $b$.

\begin{definition}
An earthquake takes as input a board and a node and outputs a board, so we write $E: \mathcal{B} \times (\mathbb{Z} \times \mathbb{Z}) \rightarrow \mathcal{B}$. If $b(s) \geq 4$, then
\begin{align*}
E[b, s](s) &= b(s)-4 \\
E[b, s](t) &= b(t)+1, \\
\end{align*}
for $ t \in n(s)$. Otherwise  $b(s) < 4$, and $E[b, s] = b$, as the earthquake does not move any chips.
We use the notation $E^n[b, \{s_k\}]$ to denote the application of $E$ to the board $b$ $n$ times, to the squares $s_1, \ldots s_{n}$ respectively. 
\end{definition}

Now that we have formalized the earthquake operator we would like to make a precise definition of stability of a board.
\begin{definition}
We say that a board is stable if all nodes contain values strictly less than 4. 
\end{definition}
This is equivalent to saying all earthquakes have no effect on the board. We would also to have a formal definition for a finite board.

\begin{definition} 
We say that a board is finite if the total number of chips on it is finite. That is a board $b$ is finite if the sum
\begin{equation}
C= \sum_{s \in \mathbb{Z} \times \mathbb{Z}} b(s), 
\end{equation}
is finite. Otherwise we say a board is infinite. 
We denote the set of finite boards as $\mathcal{B}_F$.
\end{definition}

>>>>

\begin{center}
\includegraphics[scale=0.8]{cartesiangrid.png}
\end{center}
We can coordinatize the grid graph by labeling it like the standard Cartesian grid, where we fix one node to be the origin $(0,0)$, and we index all other nodes by integer coordinates $(i,j)$. We will define the board as a function from the Cartesian grid to the nonnegative integers.

@@

This function maps node labels given as integer coordinates to chip stack sizes. For example, FIGUREREFERENCE if a board $b$ has 8 chips at the node with coordinate $(3,1)$, then $b(3,1) = 8$.

@@

We now formally define the earthquake operation. Recall that an earthquake hits a node that has 4 or more chips and moves 1 chip to each of the node's 4 neighbors.

\begin{definition}
An earthquake hits a board $b$ at node $s$, and the resulting board is denoted $E_sb$. If $b(s) \geq 4$, then
\begin{align*}
E_sb(s) &= b(s)-4,\\
E_sb(t) &= b(t)+1 \hspace{.5cm} \forall t \in n(s).\\
\end{align*}
if $b(s) < 4$, the earthquake has no effect and $E_sb = b$. 
\end{definition}

Often, a number of earthquakes will hit a board $b$. We represent the result of multiple earthquakes by a composition of the earthquake functions. For example, if an earthquake hits node 1, then node 2, and finally node 3, the resulting board is written as $E_3E_2E_1b$.

Now that we have formalized the earthquake operator, we make precise the definition of stability.
\begin{definition}
We say that a board is stable if all nodes contain values strictly less than 4. 
\end{definition}
This definition corresponds with the intuitive notion of a process that has stabilized, as stable boards can no longer be affected by earthquake operations.

Finally, we formalize the notion of $\emph{finite}$ and $\emph{infinite}$ boards, a distinction we will refer to extensively throughout this paper.

\begin{definition} 
We say that a board is finite if the total number of chips on it is finite. Otherwise we say a board is infinite. 
\end{definition}
